% Publications - a page listing research articles written using content in the thesis

\pdfbookmark[1]{Publications}{Publications} % Bookmark name visible in a PDF viewer

\chapter*{Publications} % Publications page text

%Some ideas and figures have appeared previously in the following publications:\\

%\noindent Put your publications from the thesis here. The packages \texttt{multibib} or \texttt{bibtopic} etc. can be used to handle multiple different bibliographies in your document.

%\begin{refsection}[ownpubs]
%    \small
%    \nocite{*} % is local to to the enclosing refsection
%    \printbibliography[heading=none]
%\end{refsection}

%\emph{Attention}: This requires a separate run of \texttt{bibtex} for your \texttt{refsection}, \eg, \texttt{ClassicThesis1-blx} for this file. You might also use \texttt{biber} as the backend for \texttt{biblatex}. See also \url{http://tex.stackexchange.com/questions/128196/problem-with-refsection}.

\bpar{
The following works have an highly overlapping content with this thesis:
}{
Les travaux suivants contiennent une grande partie du contenu de cette thèse:
}

\section*{Publications}


Raimbault, J. (2017). A Discrepancy-Based Framework to Compare Robustness Between Multi-attribute Evaluations. In Complex Systems Design \& Management (pp. 141-154). Springer International Publishing.

\bigskip

Raimbault, J. (2016). Investigating the Empirical Existence of Static User Equilibrium, \textit{forthcoming in EWGT 2016 proceedings, Transportation Research Procedia.} arxiv:1608.05266


\bigskip


Raimbault, J. (2016). Generation of Correlated Synthetic Data, forthcoming in \textit{Actes des Journ{\'e}es de Rochebrune 2016.}


\bigskip

Raimbault, J. (2015). Models Coupling Urban Growth and Transportation Network Growth: An Algorithmic Systematic Review Approach, forthcoming in \textit{ECTQG 2015 proceedings.} arxiv:1605.08888


\section*{Communications}


Models of growth for system of cities : Back to the simple, \textit{Conference on Complex Systems 2016, Amsterdam, Sep 2016.}



%Raimbault J., Bergeaud A. and Potiron Y. (2016). Investigating Patterns of Technological Innovation. \textit{Conference on Complex Systems 2016, Amsterdam, Sep 2016.}


\bigskip

For a Cautious Use of Big Data and Computation. \textit{Royal Geographical Society - Annual Conference 2016 - Session : Geocomputation, the Next 20 Years (1), London, Aug 2016.}


\bigskip

Indirect Bibliometrics by Complex Network Analysis. \textit{20e Anniversaire de Cybergeo, Paris, May 2016.}


\bigskip

Raimbault, J. \& Serra, H. (2016). Game-based Tools as Media to Transmit Freshwater Ecology Concepts, \textit{poster corner at SETAC 2016 (Nantes, May 2016).}


\bigskip

Le Nechet, F. \& Raimbault, J. (2015). Modeling the emergence of metropolitan transport authorithy in a polycentric urban region, \textit{ECTQG 2015, Bari, Sep 2015).}


\bigskip

Hybrid Modeling of a Bike-Sharing Transportation System, \textit{poster presented at ICCSS 2015, Helsinki, June 2015.}

\bigskip

Raimbault, J. \& Gonzales, J. (2015). Application de la Morphog{\'e}n{\`e}se de R{\'e}seaux Biologiques {\`a} la Conception Optimale d'Infrastructures de Transport, \textit{poster presented at Rencontres du Labex Dynamite, Paris, May 2015.}


