





\section[Territories and Networks]{Territories and Networks}


\subsection{Territories and Networks : There and Back Again}

\paragraph{Human Territories}

The notion of territory can be taken as a basis to explore the scope of geographical objects we will study. In Ecology, a territory corresponds to a spatial extent occupied by a group of agents or more generally an ecosystem. \emph{Human Territories} are far more complex in the sense of semiotic representations of these that are a central part in the emergence of societies. For \noun{Raffestin} in~\cite{raffestin1988reperes}, the so-called \emph{Human Territoriality} is the ``conjonction of a territorial process with an informational process'', what means that physical occupation and exploitation of space by human societies is not dissociable from the representations (cognitive and material) of these territorial processes, driving in return its further evolutions. In other words, as soon as social constructions are assumed in the constitution of human settlements, concrete and abstract social structures will play a role in the evolution of the territorial system, through e.g. propagation of information and representations, political processes, conjonction or disjonction between lived and perceived territory. Although this approach does not explicitly give the condition for the emergence of a seminal system of aggregated settlements (i.e. the emergence of cities), it insists on the role of these that become places of power and of creation of wealth through exchange. But the city has no existence without its hinterland and the territorial system can not be summarized by its cities as a system of cities. There is however compatibility on this subsystem between \noun{Raffestin} approach to territories and \noun{Pumain}'s evolutive theory of urban systems~\cite{pumain2010theorie}, in which cities are viewed as an auto-organized complex dynamical systems, and act as mediators of social changes : for example, cycles of innovation occur within cities and propagate between them. Cities are thus competitive agents that co-evolve (in the sense given before). The territorial system can be understood as a spatially organized social structure, including its concrete and abstract artifacts. A imaginary free-of-man spatial extent with potential ressources will not be a territory if not inhabited, imagined, lived, and exploited, even if the same ressources would be part of the corresponding habited territorial system. Indeed, what is considered as a ressource (natural or artificial) will depend on the corresponding society (e.g. of its practices and technological potentialities). A crucial aspect of human settlements that were studied in geography for a long time, and that relate with the previous notion of territory, are \emph{networks}. Let see how we can switch from one to the other and how their definition may be indissociable.


\paragraph{A Territorial Theory of Networks}

We paraphrase \noun{Dupuy} in~\cite{dupuy1987vers} when he proposes elements for ``a territorial theory of networks'' based on the concrete case of Urban Transportation Networks. This theory sees \emph{real networks} (i.e. concrete networks, including transportation networks) as the materialization of \emph{virtual networks}. More precisely, a territory is characterized by strong spatio-temporal discontinuities induced by the non-uniform distribution of agents and ressources. These discontinuities naturally induce a network of ``transactional projects'' that can be understood as potential interactions between elements of the territorial system (agents and/or ressources). For example today, people need to access the ressource of employments, economic exchanges operate between specialized production territories. At any time period, potential interactions existed\footnote{even when nomadism was still the rule, spatially dynamic networks of potential interactions necessarily existed, but should have less chance to materialize into concrete routes.% bib on that ?
}. The potential interaction network is concretized as offer adapts to demand, and results of the combination of economic and geographical constraints with demand patterns, in a non-linear way through agents designed as \emph{operators}. This process is not immediate, leading to strong non-stationarity and path-dependance effects : the extension of an existing network will depend on previous configuration, and depending on involved time scales, the logic and even the nature of operators may have evolved. \noun{Raffestin} points out in his preface of~\cite{offner1996reseaux} that a geographical theory articulating space, network and territories had never been consistently formulated. It appears to still be the case today, but the theory developed just before is a good candidate, even if it stays at a conceptual level. The presence of a human territory necessarily imply the presence of abstract interaction networks and concrete networks used for transportation of people and ressources (including communication networks as information is a crucial ressource). Depending on regime in which the considered system is, the respective role of different networks may be radically different. Following \noun{Duranton} in \cite{duranton1999distance}, pre-industrial cities were limited in growth because of limitations of transportation networks. Technological progresses have lead to the end of these limitations and the preponderance of land markets in shaping cities (and thus a role of transportation network as shaping prices through accessibility), and recently to the rising importance of telecommunication networks that induce a ``tyranny of proximity'' as physical presence is not replaceable by virtual communication. This territorial approach to networks seems natural in geography, since networks are studied conjointly with geographical objects with an underlying theory, in opposition to network science that studies brutally spatial networks with few thematic background~\cite{ducruet2014spatial}.


\paragraph{Networks shaping territories ?}

% how do network shape territories : boundaries, scales, etc.
% example : \cite{l2012ville} bahn-ville, volontary coevol ? // idem villes nouvelles

However networks are not only a material manifestation of territorial processes, but play their part in these processes as they evolution may shape territories in return. In the case of \emph{technical networks}, an other designation of real networks given in~\cite{offner1996reseaux}, many examples of such feedbacks can be found : the interconnectivity of transportation networks allows multi-scalar mobility patterns, thus shaping the lived territory. At a smaller scale, changes in accessibility may result in an adaptation of a functional urban space. Here emerges again an intrinsic difficulty : it is far from evident to attribute territorial mutations to some network evolutions and reciprocally materialization of a network to precise territorial dynamics. Coming back to Diderot should help, in the sense that one must not consider network nor territories as independent systems that would have causal relationships but as strongly coupled components of a larger system. The confusion on possible simple causal relationships has fed a scientific debate that is still active. Methodologies to identify so-called \emph{structural effects} of transportation networks were proposed by planners in the seventies~\cite{bonnafous1974detection,bonnafous1974methodologies}. It took some time for a critical positioning on unreasoned and decontextualized use of these methods by planners and politics generally to technocratically justify transportation projects, that was first done by \noun{Offner} in~\cite{offner1993effets}. Recently the special issue~\cite{espacegeo2014effets} on that debate recalled that on the one hand misconceptions and misuses were still greatly present in operational and planning milieus as~\cite{crozet:halshs-01094554} confirmed, and on the other hand that a lot of scientific progresses still need to be made to understand relations between networks and territories as \noun{Pumain} highlights that recent works gave evidence of systematic effects on very long time scales (as e.g. the work of \noun{Bretagnolle} on railway evolution, that shows a kind of structural effect in the necessity of connectivity to the network for cities to ``stay in the game'', but that is not fully causal as not sufficient). At a macroscopic level typical patterns of interaction emerge, but microscopic trajectories of the system are essentially chaotic : the understanding of coupled dynamics strongly depends on the scale considered. At a small scale it seems indeed impossible to show systematic behavior, as \noun{Offner} pointed out. For example, on comparable French mountain territories, \cite{berne2008ouverture} shows that reactions to a same context of evolution of the transportation network can lead to very different reactions of territories, some finding a huge benefit in the new connectivity, whereas others become more closed. These potential retroactions of networks on territories does not necessarily act on concrete components : \noun{Claval} shows in~\cite{claval1987reseaux} that transportation and communication networks contribute to the collective representation of territories by acting on territorial belonging feeling.




\paragraph{Territorial Systems}

This detour from territories, to networks and back again, allows us to give a preliminary definition of a territorial system that will be the basis of our following theoretical considerations. As we emphasized the role of networks, the definition takes it into account.

\bigskip

\textbf{Preliminary Definition.} \textit{A territorial system is a human territory to which both interaction and real networks can be associated. Real networks are a component of the system, involved in evolution processes, through multiples feedbacks with other components at various spatial and temporal scales.}


\bigskip


This reading of territorial systems is conditional to the existence of networks and may discard some human territories, but it is a deliberate choice that we justify by previous considerations, and that drives our subject towards the study of interactions between networks and territories.



\subsection{Transportation Networks}


\paragraph{The particularity of transportation networks}

Already evoked in relation to the question of structural effects of networks, transportation networks play a determining role in the evolution of territories. Although other types of networks are also strongly involved in the evolution of territorial systems (see e.g. the discussions of impacts of communication networks on economic activities), transportation networks shape many other networks (logistics, commercial exchanges, social concrete interactions to give a few) and are prominent in territorial evolution patterns, especially in our recent societies that has become dependent of transportation networks~\cite{bavoux2005geographie}. The development of French High Speed Rail network is a good illustration of the impact of transportation networks on territorial development policies. Presented as a new era of railway transportation, a top-down planning of totally novel lines was introduced as central for developments~\cite{zembri1997fondements}. The lack of integration of these new networks with existing ones and with local territories is now observed as a structural weakness and negative impacts on some territories have been shown~\cite{zembri2008contribution}. A review done in~\cite{bazin2011grande} confirms that no general conclusions on local effects of High Speed lines connection can be drawn although it keeps a strong place in imaginaries. These are examples of how transportation networks have both direct and indirect impacts on territorial dynamics. Integrated planning, in the sense of a joint planning of transportation infrastructures and urban development, considers the network as a determining component of the territorial system. Parisian \emph{Villes Nouvelles} are such a case, that witnesses of the complexity of such planning actions that generally do not lead to the desired effect~\cite{es119}. Recent projects as~\cite{l2012ville} have try to implement similar ideas but we have now not enough temporal scope to judge their success in effectively producing an integrated territory. Transportation networks are anyway at the center of these approaches of urban territories. We will focus in our work on transportation networks for the various reasons given here.




\paragraph{Deconstructing Accessibility}

% critic of accessibility as a planning tool : danger of not taking into account socio-eco dynamics and coupled dynamics (coevol) - cit Hadri mobility as a constructed notion.

The notion of accessibility comes rapidly when considering transportation networks. Based on the possibility to access a place through a transportation network (including transportation speed, difficulty of travel), it is generally described as a potential of spatial interaction\footnote{and often generalized as \emph{functional accessibility}, for example employments accessible for actives at a location. Spatial interaction potentials ruling gravity law can also been understood this way.}~\cite{bavoux2005geographie}. This object is often used as a planning tool or as an explicative variable of agents localisation for example. We must warn here on the potential dangers of its unconditional use. More precisely, it may be a construction that misses a consistent part of territorial dynamics. The mystification of the notion of \emph{mobility} was shown by \noun{Commenges} in~\cite{commenges:tel-00923682}, which proved than most of debates on modeling mobility and corresponding notions were mostly made-of by transportation administrators of \emph{Corps des Ponts} who roughly imported ideas from the United States without adaptation and reflexion fit to the totally different French context. Accessibility may be such a social construct and have no theoretical root since it is mostly a modeling and planning tool. Recent debates on the planification of \emph{Grand Paris Express}~\cite{confMangin}, a totally novel metropolitan transportation infrastructure planned to be built in the next twenty years, have revealed the opposition between a vision of accessibility as a right for disadvantaged territories against accessibility as a driver of economic development for already dynamic areas, both being difficultly compatible since corresponding to very different transportation corridors. Such operational issues confirm the complexity of the role of transportation networks in the dynamics of territorial systems, and we shall give in our work elements of response to a definition of accessibility that would integrate intrinsic territorial dynamics.


\paragraph{Scales and Hierarchies}


% \cite{10.1371/journal.pone.0102007}
% \cite{Tsekeris20131} : congestion related to land-use

An incontournable aspect of transportation networks that we will need to take into account in further developments is hierarchy. Transportation networks are by essence hierarchical, depending on scales they are embedded in. \cite{10.1371/journal.pone.0102007} showed empirical scaling properties for public transportation networks for a consequent number of metropolitan areas across the world, and scaling laws reveal the presence of hierarchy within a system, as for size hierarchy for system of cities expressed by Zipf's law~\cite{nitsch2005zipf} or other urban scaling laws\cite{2013arXiv1301.1674A,2015arXiv151000902B}. Transportation network topology has been shown to exhibit such scaling also for the distribution of its local measures such as centrality~\cite{samaniego2008cities}. Hierarchy seems to play a particular role on interaction processes, as \noun{Bretagnolle}~\cite{bretagnolle:tel-00459720} highlighted an increasing correlation between urban hierarchy and network hierarchy for French railway network, marker of positive feedbacks between urban rank and network centralities. Different regimes in space and times were identified : for French railway network evolution e.g., a first phase of adaptation of the network to the existing urban configuration was followed by a phase of co-evolution i.e. in the sense that causal relations become difficult to identify. Railway evolution in the United States followed a different pattern, without hierarchical diffusion, shaping locally urban growth. It emphasizes the presence of path-dependance for trajectories of urban systems : the presence in France of a previous city system and network (postal roads) strongly shaped railway development, whereas its absence in the US lead to completely different dynamics. An open question is if generic processes underlie both evolutions, each being different realizations with different initial conditions and different meta-parameters (different \emph{regimes} in the sense of settlement systems transitions introduced in the current ANR Research project TransMonDyn, as a transition can be understood as a change of stationarity for meta-parameters of a general dynamic). In terms of dynamical systems formulation, it is equivalent to ask if dynamics of attractors (long time scale components) obey similar equations as the position and nature of attractors for a stochastic dynamical system give its current regime, in particular if it is in a divergent state (positive local Liapounov exponent) or is converging towards stable mechanisms~\cite{sanders1992systeme}. To answer this question together with a disentangling of co-evolution processes for that regime, \cite{bretagnolle:tel-00459720} proposes modeling as a constructive element of answer. We will see in next section how modeling can bring knowledge about territorial processes.



\paragraph{Interactions between transportation networks and territory}

At this state of progress, we have naturally identified a research subject that seems to take a significant place in the complexity of territorial systems, that is the study of interactions between transportation networks and territories. In the frame of our preliminary definition of a territorial system, this question can be reformulated as the study of networked territorial systems with an emphasize on the role of transportation networks in system evolution processes.




%----------------------------------------------------------------------------------------

\newpage

\section{Modeling Interactions}


\subsection{Modeling in Quantitative Geography}

% brief reference to the history of TQG ; history of modeling.
%  note : history of future of TQG, London september 2016

Modeling in Theoretical and Quantitative Geography (TQG), and more generally in Social Science, has a long history on which we can not go further than a general context. \noun{Cuyala} does in~\cite{cuyala2014analyse} an analysis of the spatio-temporal development of French speaking TQG movement and underlines the emergence of the discipline as the combination between quantitative analysis (e.g. spatial analysis or modeling and simulation practices) and theoretical constructions, an integration of both allowing the construction of theories from empirical facts that yield theoretical hypothesis to be tested on empirical data. These approach were born under the influence of the \emph{new geography} in Anglo-saxon countries and Sweden. A broad history of the genesis of models of simulation in geography is done by \noun{Rey} in~\cite{rey2015plateforme} with a particular emphasis on the notion of validation of models. The use of computation for simulation of models is anterior to the introduction of paradigms of complexity, coming back to \noun{H{\"a}gerstrand} and \noun{Forrester}, pioneers of spatial economic models inspired by Cybernetics. With the increase of computational possibilities epistemological transformations have also occurred, with the apparition of explicative models as experimental tools. \noun{Rey} compares the dynamism of seventies when computation centers were opened to geographers to the democratization of High Performance Computing (transparent grid computing, see \cite{schmitt2014half} for an exemple of the possibilities offered in terms of model validation and calibration, decreasing the computational time from 30 years to one week), that is also accompanied by an evolution of modeling practices~\cite{banos2013pour} and techniques~\cite{10.1371/journal.pone.0138212}. Modeling (in particular computational models of simulation) is seen by many as a fundamental building brick of knowledge : \cite{livet2010} recalls the combination of empirical, conceptual (theoretical) and modeling domains with constructive feedbacks between each. A model can be an exploration tool to test assumptions, an empirical tool to validate a theory against datasets, an explicative tool to reveal causalities (and thus internal processes of a system), a constructive tool to iteratively build a theory with an iterative construction of an associated model. These are example among others : \noun{Varenne} proposes in~\cite{varenne2010simulations} a refined classifications of diverse functions of a model. We will consider modeling as a fundamental instrument of knowledge on processes within complex adaptive systems, as already evoked, and restraining again our question, will focus on \emph{models involving interactions between transportation networks and territories}.



\subsection{Modeling Territories and Networks}

% here overview of different approaches
% Q : do it here, not during quant epistemo part ?

Concerning our precise question of interactions between transportation networks and territories, we propose an overview of existing approaches. Following~\cite{bretagnolle2002time}, the ``\textit{thoughts of specialists in planning aimed to give definitions of city systems, since 1830, are closely linked to the historical transformations of communication networks}''. It implies that ontologies and corresponding models addressed by geographers and planners are closely linked to their current historical preoccupations, thus necessarily limited in scope and purpose. In a perspectivist vision of science~\cite{giere2010scientific} such boundaries are the essence of the scientific entreprise, and as we will argue in chapter~\ref{ch:theory} their combination and coupling in the case of models is a source of knowledge.


\subsubsection{Land-Use Transportation Interaction Models}

A subsequent bunch of literature in modeling interaction between networks and territories can be found in the field of planning, with the so-called \emph{Land-use Transportation Interaction Models}. These works are difficult to be precisely bounded as they may be influenced by various disciplines. For example, from the point of view of Urban Economics, propositions for synthesizing models have existed for a relatively long term~\cite{putman1975urban}. The variety of possible models has lead to operational comparisons~\cite{paulley1991overview,wegener1991one}. More recently, the respective advantages of static and dynamic modeling was investigated in~\cite{kryvobokov2013comparison}. Generally these type of models operate at relatively small temporal and spatial scales. \cite{wegener2004land} reviewed state of the art in empirical and modeling studies on interactions between land-use and transportation. It is positioned in economic, planning and sociological theoretical contexts, and is relatively far from our geographical approach aiming to understand long-time processes. Seventeen models are compared and classified, none of which implements actually network endogenous evolution on the relatively small time scales of simulation. A complementary review done in \cite{chang2006models} broadens the scope with inclusion of more general classes of models, such as spatial interaction models (including traffic assignment and four steps models), operational research planning models (optimal localisations), micro-based random utility models, and urban market models. These techniques operate also at small scales and consider at most land-use evolution. \cite{iacono2008models} covers a similar scope with a further emphasis on cellular automata models of land-use change and agent-based models. These type of models are still largely developed and used today, as for example \cite{delons:hal-00319087} which is used for Parisian metropolitan region. The short-term range of application and their operational character makes them useful for planning, what is far from our preoccupation to obtain explicative models for geographical processes. 





\subsubsection{Network Growth}

% economic models
%\cite{yerra2005emergence} % : cost-driven model of nw reinforcement (// slime mould)
%\cite{louf2013emergence} % : trade-off between cost and benefits due to flows -> compare with lutecia rules ?
%\cite{xie2009modeling} review of network growth economic appraoches
%\cite{bigotte2010integrated} % : planning network - similar to nw growth ? - hierarchy in cities and nw
 
 % geometric - local optimization
%\cite{barthelemy2008modeling}
%\cite{courtat2011mathematics} % measures and morphogenesis. Vision of morphogenesis as living organism : nuance that in theory.
%\cite{de2007netlogo} % geom rules in Tijuana model
%\cite{rui2013exploring} % local based optimisation morphogenesis model. RQ : quote only physicists work -> justification for extended quantitative epistemology ?
%\cite{yamins2003growing} strange model
 
% biological nws

%\cite{tero2010rules} % physarum : biological nw heuristics
%\cite{tero2006physarum} % potentialities of physarum machines, here for routing.
%\cite{adamatzky2010road} % planning absurdities
%\cite{zhu2013amoeba} % TSP solving : long range correlations.

Network growth can be used to design modeling entreprises that aim to endogenously explain growth of transportation networks, generally from a bottom-up point of view, i.e. by exhibiting local rules that would allow to reproduce network growth over long time scales (generally the road network). Economists have proposed such models : \cite{zhang2007economics} reviews transportation economics literature on network growth within an endogenous growth theory~\cite{aghion1998endogenous}, recalling the three main features studied by economists on that subject that are road pricing, infrastructure investment and ownership regime, and describes an analytical model combining the three. \cite{xie2009modeling} develops a broad review on network growth modeling extending to other fields : transportation geography early developed empirical-based models but which did concentrate on topology reproduction rather than on mechanisms according to~\cite{xie2009modeling} ; statistical models on case studies provide mitigated conclusions on causal relations between offer and demand ; economists have studied infrastructure provision from both microscopic and macroscopic point of views, generally non-spatial ; network science has provided toy-models of network growth based on structural and topological rules rather on mechanism-based rules. An other approach not mentioned that we will develop further is biologically inspired network design. We first give some example of economic-based and geometrical-based network growth modeling attempts. \cite{yerra2005emergence} shows through a reinforcement economic model including investment rule based on traffic assignment that local rules are enough to make hierarchy of roads emerge for a fixed land-use. A very similar model in~\cite{louf2013emergence} with simpler cost-benefits obtains the same conclusion. Whereas these models based on processes focus on reproducing macroscopic patterns of networks (typically scaling), geometrical optimization models aim to ressemble topologically real networks. \cite{barthelemy2008modeling} proposes a model based on local energy optimization but it stays very abstract and unvalidated. The morphogenesis model given in~\cite{courtat2011mathematics} based on local potential and connectivity rules, even if not calibrated, seems to reproduce more reasonably real street patterns. Very close work is done in~\cite{rui2013exploring}. Other tentatives \cite{de2007netlogo,yamins2003growing} are closer to procedural modeling~\cite{lechner2004procedural,watson2008procedural} and therefore not of interest in our purpose as they can difficultly be used as explicative models. Finally, an interesting and original approach to network growth are biological networks. These belong to the field of morphogenetic engineering pioneered by \noun{Doursat} that aim to design artificial complex system inspired from natural complex systems and in which a control of emerging properties is possible~\cite{doursat2012morphogenetic}. \emph{Physarum Machines}, that are models of a self-organized mould (slime mould) have been shown to provide efficient bottom-up solution to computationally heavy problems such as routing problems~\cite{tero2006physarum} or NP-complete navigation problems such as the Travelling Salesman Problem~\cite{zhu2013amoeba}. It has been shown to produce networks with Pareto-efficient cost-robustness properties~\cite{tero2010rules}, relatively close in shape to real networks (under certain conditions, see~\cite{adamatzky2010road}). This type of models can be of interest for us since auto-reinforcement mechanisms based on flows are analog to mechanisms of link reinforcement in transportation economics.
% note : comparison with Francois model for french railway (nothing published yet)




\subsubsection{Hybrid Modeling}



%\cite{bigotte2010integrated}
%\cite{levinson2007co} : economic model of coevolution. % check timescales if are consistent.
%\cite{levinson2005paving} % markov chain : not really modeling but more statistics.
%\cite{raimbault2014hybrid}



Models of simulation implementing a coupled dynamic between urban growth and transportation network growth are relatively rare, and always rather poor from a theoretical and thematic point of view. A generalization of the geometrical local optimization model described before was developed in~\cite{barthelemy2009co}. % pb of scales, def of coevolution, thematic meaning of assumptions, etc.
As for the road growth model of which it is an extension, no thematic nor theoretical justification of local mechanisms is provided, and the model is furthermore not explored and no geographical knowledge can be drawn from it. \cite{levinson2007co} adopts a more interesting economic approach, similar to a four step model (gravity-based origin-destination flows generation, stochastic user equilibrium traffic assignment) including travel cost and congestion, coupled with a road investment module simulating toll revenues for constructing agents, and a land-use evolution module updating actives and employments through discrete choice modeling. The experiments showed that co-evolving network and land uses lead to positive feedbacks reinforcing hierarchy, but are far from satisfying for two reasons : first network topology does not really evolve as only capacities and flows change within the network, what means that more complex mechanisms on longer time scales are not taken into account, and secondly the conclusions are very limited as model behavior is not known since sensitivity analysis is done on few one-dimensional spaces : exhaustive mechanisms stay thus unrevealed as only particular cases are described in the sensitivity analysis. From an other point of view, \cite{levinson2005paving} is also presented as a model of co-evolution, but corresponds more to coupled statistical analysis as it relies on a Markov-chain predictive model. \cite{rui2011urban} gives a model in which coupling between land-use and network growth is done in a weak paradigm, land-use and accessibility having no feedback on network topology evolution. \cite{achibet2014model} describes a co-evolution model at a very small scale (scale of the building), in which evolution of both network and buildings are ruled by a same agent (influenced differently by network topology and population density) what implies a too strong simplification of underlying processes. Finally, a simple hybrid model explored and applied to a toy planning example in~\cite{raimbault2014hybrid}, relies on urban activities accessibility mechanisms for settlement growth with a network adapting to urban shape. The rules for network growth are too simple to capture processes we are interested in, but the model produces at a small scale a broad range of urban shapes reproducing typical patterns of human settlements.


\subsubsection{Urban Systems Modeling}

% differentiate to other hybrid models : SimpopNet and others ? (if exist ?)

An approach closer to our current questioning is the one of integrated modeling of system of cities. In the continuity of Simpop models for city systems modeling, \noun{Schmitt} described in~\cite{schmitt2014modelisation} the SimpopNet model which aim was precisely to integrate co-evolution processes in system of cities on long time scales, typically rules for hierarchical network development as a function of cities dynamics coupled with city dynamics depending on network topology. Unfortunately the model was not explored nor further studied, and furthermore stayed at a toy-level. \noun{Cottineau} proposed transportation network endogenous growth as the last building bricks of her Marius productions but it stayed at a conceptual construction stage. We shall position more in that stream of research in this thesis.



\subsection{Sketch of a \emph{Modelography}}

An ongoing work is the production of a synthesis of this overview, from a modular modeling point of view, combined with a purpose and scale classification. Already mentioned, modular modeling consists in the integration of heterogeneous processes and implementation of processes in order to extract the set of mechanisms giving the best fit to empirical data~\cite{cottineau2015incremental}. We can thus classify models described here according to their building bricks in terms of processes implemented and thus identify possible coupling potentialities. This work is a preliminary step for the analysis in quantitative epistemology developed in chapter~\ref{ch:quantepistemo}.



%-------------------------


\section{Research Question}

To close this thematic touring introducing chapter, we can state a general research question that frames our further theoretical constructions and first modeling attempts. It is roughly the same as the problematic given at the end of previous section, but adding the insight of modeling as the approach to understand these complex systems.

%networked territorial systems with an emphasize on the role of transportation networks in system evolution processes.

\medskip

\textbf{General research Question.} \textit{To what extent a modeling approach to territorial systems as networked human territories can help disentangling complexly involved processes ?}

\medskip

This question will be refined by theoretical developments in the next chapter and experiments in the followings.









Theory is a key element of any scientific construction, especially in Human Sciences in which object definition and questioning are more open but also determining for research directions. We develop in this chapter a self-consistent theoretical background. It naturally emerges from thematic considerations of previous chapter, empirical explorations done in chapter~\ref{ch:empirical} and modeling experiments conducted in chapter~\ref{ch:modeling}, as a linear structure of knowledge is not appropriate to translate the type of scientific entreprise we are conducting, typically in the spirit of \noun{Sanders} in~\cite{livet2010} for which the simultaneous conjonction of empirical, conceptual and modeling domains is necessary for the emergence of knowledge. This theoretical construction is however presented to be understood independently, and is used as a structuring skeleton for the rest of the thesis.

We propose first to construct the \emph{geographical theory} that will pose the studied objects and their meaning in the real world (their ontology), with their interrelations. This yields precise assumptions that will be sought to be confirmed or proven false in the following. Staying at a thematic level appears however to be not enough to obtain general guidelines on the type of methodologies and the approaches to use. More precisely, even if some theories imply an more natural use of some tools\footnote{to give a rough example, a theory emphasizing the complexity of relations between agents in a system will conduct generally to use agent-based modeling and simulation tools, whereas a theory based on macroscopic equilibrium will favorise the use of exact mathematical derivations.}, at the subtler level of contextualization in the sense of the approach taken to implement the theory (as models or empirical analysis), the freedom of choice may mislead into unappropriated techniques or questionings (see \cite{raimbault2016cautious} on the example of incautious use of big data and computation). We develop therefore in a second section a theoretical framework at a meta-level, aiming to give a vision and framing for modeling socio-technical systems.



%----------------------------------------------------------------------------------------

\newpage

% first section that develops elements of geographical theory : system of cities, territories, etc.
%  city morphogenesis ? -> link to other approaches of morphogenesis (Turing)
%  coevolution : give a precise theoretical meaning

% HERE precise exact definition, framed precisely (reference for the following).

\section{Geographical Theoretical Context}



%%%%%%%%%%%
\subsection{Foundation}



%%%%%%%%%%%%%%
\subsubsection{Networked Human Territories}

Our first pillar has already been constructed before in the thematic exploration of the research subject. We rely on the notion of \emph{Human Territory} elaborated by \noun{Raffestin} as the basis for a definition of territorial systems. It permits to capture complex human geographical systems in their concrete and abstract characteristics and representation. For example, a metropolitan territorial system can be apprehended simply by the functional extent of daily commuting, or by the perceived or lived space of different populations, the choice depending on the precise question asked. Note that this approach to territory is a position and that other (possibly compatible) entries could be taken~\cite{murphy2012entente}. The concrete of this pillar in reinforced by the territorial theory of networks of \noun{Dupuy}, yielding the notion of networked human territory, as a human territory in which a set of potential transactional networks have been realized, which is in accordance with vision of the territory as networked places~\cite{champollion:halshs-00999026}. We make therein the assumption that real networks are necessary elements of territorial systems.


%%%%%%%%%%%%%%%
\subsubsection{Evolutive Urban Theory}

% development of Denise theory
%  -> extension with precision on coevolution ? (read Holland on coevolution) -- beware of biological //

The second pillar of our theoretical construction is the Evolutive Urban Theory of \noun{Pumain}, closely linked to the complexity approach we take. This theory was first introduced in~\cite{pumain1997pour} which argues for a dynamical vision of city systems, in which self-organization is key. Cities are interdependent evolutive spatial entities whose interrelations produces the macroscopic behavior at the scale of city system. The city system is also designed as a network of city what emphasizes its view as a complex system. Each city is itself a complex system in the spirit of~\cite{berry1964cities}, the multi-scale aspect being essential in this theory, since microscopic agents convey system evolution through complex feedbacks between scales. The positioning within Complex System Sciences was later confirmed~\cite{pumain2003approche}. It was shown that this theory provide an interpretation for the origin of pervasive scaling laws, resulting from the diffusion of innovation cycles between cities~\cite{pumain2006evolutionary}. The aspect of resilience of system of cities, induced by the adaptive character of these complex systems, implies that cities are drivers and adapters of social change~\cite{pumain2010theorie}. Finally, path dependance yield non-ergodicity within these systems, making ``universal'' interpretations of scaling laws developed by physicists incompatible with evolutive urban theory~\cite{pumain2012urban}. We will interpret territorial systems following that idea of complex adaptive systems.




%%%%%%%%%%%
\subsubsection{Urban Morphogenesis}

% -> make a link between city systems and urban form/cityscape / territorial configurations

% Why morphogenesis is important : linked with modularity and scale -> if a submodule can be explained independantly (ie morphigeneis process is isolated), then we have the characteristic scale. then when size grows and interaction within city system -> can not explain alone (or with externalities ?) -> need a change in scale. ex. influecne of city system for size, activities; posiiton of an airport in a metropoltian region ; emergence of MCR.  ==> Assimptions to be tested with models ?.

% Alexander and Salingaros

% include transportation network, hierarchy and congestion in transport : Remy vs Benjamin (paper ? -> see with René)


The idea of morphogenesis was introduced by \noun{Turing} in~\cite{turing1952chemical} when trying to isolate simple chemical rules that could lead to the emergence of the embryo and its form. The morphogenesis of a system consists in self-consistent evolution rules that produce the emergence of its successives states, i.e. the precise definition of self-organization. Progresses towards the understanding of embryo morphogenesis (in particular the isolation of processes producing the differentiation of cells from an unique cell) has been made only recently with the use of Complexity Approaches in integrative biology~\cite{delile2016chapitre}. In the case of urban systems, the idea of urban morphogenesis, i.e. of self-consistent mechanisms that would produce the urban form, is more used in the field of architecture and urban design~\cite{hachi2013master} (as \noun{Alexander} generative grammar ``Pattern Language'' e.g.), in relation with theories of Urban Form~\cite{moudon1997urban}. This idea can be pushed into very small scales such as the building~\cite{whitehand1999urban} but we will use it more at a mesoscopic scale, in terms of land-use changes within an intermediate scale territorial system, in the same ontologies as Urban morphogenesis modeling literature (for example \cite{bonin2012modele} describes a model of urban morphogenesis with qualitative differentiation, whereas \cite{makse1998modeling} give a model of urban growth based on a mono-centric population distribution perturbed with correlated noises). The notion of morphogenesis will be important in our theory in link with modularity and scale. Modularity of a complex system consists in its decomposition into relatively independent sub-modules, and modular decomposition of a system can be seen as a way to disentangle non-intrinsic correlations~\cite{2015arXiv150904386K} (think of a block diagonalisation of a first order dynamical system). The isolation of a subsystem yields a corresponding characteristic scale. Isolating possible morphogenesis processes imply a controlled isolation (controlled boundary conditions e.g.) of the considered system, corresponding to a modularity level and thus a scale. When self-consistent processes are not enough to explain the evolution of the system (with reasonable action on boundary conditions), a change of scale is necessary, caused by an underlying phase transition in modularity. The example of metropolitan growth is a good example : complexity of interactions within the metropolitan region will grow with size and diversity of functions leading to a change in scale necessary to understand processes. The emergence of an international airport will strongly influence local development, what corresponds to the significant integration within a larger system. The characteristic scales and processes for which these change occur will be precise questions to be investigated through modeling. It is interesting to remark that a territorial subsystem in which morphogenesis has a sense can be seen as an \emph{autopoietic system} in the extended sense of \noun{Bourgine} in~\cite{bourgine2004autopoiesis}, as a network of auto-reproducing processes\footnote{which are however not cognitive, making this auto-organized systems fortunately not alive in the sense of autopoietic and cognitive systems} regulating their boundary conditions, what emphasizes boundaries on which we will last insist.


% transition : Bourgine autopoiesis -> importance of boundaries -> link to Holland.

%%%%%%%%%%%
\subsubsection{Co-evolution}

% other insight : Holland Signal and Boundaries, ecological niche etc. : contextualize within this framework, clarify definition of co-evolution

Our last pillar is a clarification of the notion of \emph{co-evolution}, on which \noun{Holland} shed light through an approach of complex adaptive systems by a theory of CAS as signal processing agents operating thanks to their boundaries~\cite{holland2012signals}. In this theory, complex adaptive systems form aggregates at diverse hierarchical levels, that correspond to different level of self-organization, and boundaries are vertically and horizontally intricate in a complex way. That approach introduces the notion of \emph{niche} as a relatively independent subsystem in which ressources circulate (the same way as network communities) : numerous illustrations are given such as economical niches or ecological niches. Agents within a niche are said to be \emph{co-evolving}. Co-evolution thus means strong interdependences (implying circular causal processes) and a certain independence regarding the exterior of the niche. The notion is naturally flexible as it will depend on ontologies, resolution, thresholds etc. considered to define the system. This concept is easily transmissible to the evolutive urban theory and converges with the notion of co-evolution described by \noun{Pumain} : co-evolving agents in a system of cities consist in a niche with its flows, signals and boundaries and thus co-evolving entities in the sense of \noun{Holland}. This notion will be important for us in the definition of territorial subsystems and their coupling.



%%%%%%%%%%%
%\subsection{Requirements}
% RQ : no requirements for the theory, contained within pillars : requirement is the presence of these pillars ?



%%%%%%%%%%%
\subsection{Synthesis : an theory of co-evolutive networked territorial systems}

% put different elements together and construct the geographical theory
% give here precise definitions

We synthesize our pillars as a short self-consistent geographical theory of territorial systems in which networks play a central role in the co-evolution of components of the system. See the foundation subsection for definitions and references. The formulation is intended to be minimalistic.

\medskip

\begin{definition}
\textbf{ - Territorial System.} A territorial system is a set of networked human territories, i.e. human territories in and between which real networks exist.
\end{definition}

\medskip

At this step complexity and dynamical evolutive characters of territorial systems are implied but not an explicit part of the theory. We will assume to simplify a discrete definition of temporal, spatial and ontological scales under modularity and local stationarity assumptions.

% definition of scale and stationarity
%\textit{Equivalence between existence of discrete scales and discrete stationarity levels ?}

\medskip

\begin{proposition}
\textbf{ - Discrete scales.} Assuming a discrete modular decomposition of a territorial system, the existence of a discrete set $(\tau_i,x_i)$ of temporal and functional scales for the territorial system is equivalent to the local temporal stationarity of a random dynamical system specification of the system.
\end{proposition}

\begin{proof}
\textbf{(Sketch of).} We underlie that any territorial system can be represented by random variables, what is equivalent to have well defined objects and states and use the Transfer Theorem on events of successive states. If $X=(X_j)$ is the modular decomposition, we have necessarily quasi-independence of components in the sense that $\Covb{dX_j}{dX_{j'}}\simeq 0$ at any time. General stationarity transitions induce modular transitions that are kept or not depending if they correspond to an effective transition within the subsystem, what provide temporal scales as characteristic times of sub-dynamics. Functional scales are the corresponding extent in the state space.\qed
\end{proof}

% assumption : existence of scales

\medskip

This proposition induce a discrete representation of system dynamics in time. Note that even in the case of no modular representation, the system as a whole will verify the property. This definition of scales allows to explicitly introduce feedback loops and thus emergence and complexity, making our theory compatible with the evolutive urban theory.



\begin{assumption}
\textbf{ - Scales and Subsystems intrication. } Complex networks of feedbacks exist both between and inside scales, what impose the existence of weak emergence~\cite{bedau2002downward}. Furthermore a horizontal and vertical hierarchical imbrication of boundaries is not the rule.
\end{assumption}

% co-evolution

Within these complex subsystems intrications we can isolate co-evolving components using morphogenesis. The following proposition is a consequence of the equivalence between the independence of a niche and its morphogenesis. Morphogenesis provides the modular decomposition (local stationarity assumed) needed for the existence of scale, giving minimal vertically (scale) and horizontally (space) independent subsystems.

\begin{proposition}
\textbf{ - Co-evolution of components. } Morphogenesis processes of a territorial system are an equivalent formulation of the existence of co-evolutive subsystems.
\end{proposition}


% importance of nws as necessary subcomponents
%  maybe where we diverge from Denise theory ?

Finally we make a key assumption putting real networks at the center of co-evolutive dynamics, introducing their necessity to explain dynamical processes of territorial systems.

\begin{assumption}
\textbf{ - Necessity of Networks. } Network evolution can not be explained only by the dynamics of other territorial components and reciprocally, i.e. co-evolving territorial subsystems include real networks. They can thus be at the origin of regime changes (transition between stationarity regimes) or more dramatic bifurcations in dynamics of the whole territorial system.
\end{assumption}

On long time scale, an overall co-evolution has been shown for the french railway network by~\cite{bretagnolle:tel-00459720}. At smaller scales it is less evident (debate on structural effects) but we postulate that co-evolution effects are present at any scale. Regional examples may illustrate that : Lyon has not the same dynamical relations with Clermont than with Saint-Etienne and network connectivity has necessarily a role in that (among intrinsic interaction dynamics and distance). At a smaller scale, we think that effects are even less observable, but precisely because of the fact that co-evolution is stronger and local bifurcations will occur with stronger amplitude ans greater frequency than in macroscopic systems where attractors are more stable and stationarity scales greater. We will try to identify bifurcation or phase transitions in toy models, hybrid models and empirical analysis, at different scales, on different case studies and with different ontologies.

One difficulty in our construction is the stationarity assumption. Even if it seems a reasonable assumptions on large scales and has already been observed in empirical data~\cite{sanders1992systeme}, we shall verify it in our empirical studies. Indeed, this question is at the center of current research efforts to apply deep learning techniques to geographical systems : \noun{Bourgine} has recently developed a framework to extract patterns of Complex Adaptive Systems\footnote{Using a representation theorem~\cite{knight1975predictive}, any discrete stationary process is a \emph{Hidden Markov Model}. Given the definition of a causal state as $\Pb{future | A} = \Pb{future | B}$, the partition of system states induced by the corresponding equivalence relations allows to derive a \emph{Recurrent Network} that is enough to determine next state of the system, as it is a \emph{deterministic} function of previous state and hidden states~\cite{shalizi2001computational} : $(x_{t+1},s_{t+1}) = F\left[(x_t,s_t)\right]$. The estimation of Hidden States and of the Recurrent Function thus captures through deep learning entirely dynamical patterns of the system, i.e. full information on its dynamics and internal processes.}. The issues are then if the stationarity assumption be tackled through augmentation of system states, and if heterogeneous and asynchronous data be used to bootstrap long time-series necessary for a correct estimation of the neural network. These issue are related to the stationarity assumption for the first and to non-ergodicity for the second.







