\input{header.tex}


\title{Towards a Theory of Co-evolutive Networked Territorial Systems: Insights from Transportation Governance Modeling in Pearl River Delta, China\\\bigskip
\bigskip
\bigskip
\textit{Medium Project Seminar, December 2016}
}\bigskip
\bigskip
\author{\noun{Juste Raimbault}$^{1,2}$\\
\small(1) UMR CNRS 8504 Géographie-cités and (2) UMR-T IFSTTAR 9403 LVMT
}
%\date{15 octobre 2015}
\date{}

\maketitle

\justify

\pagenumbering{gobble}



\vspace{0.2cm}

\textbf{Keywords : }\textit{Co-evolutive Networked Territorial Systems ; Agent-based Modeling ; Transportation Governance ; Pearl River Delta Mega-city Region}

\vspace{0.5cm}

%
% Theoretical and Quantitative Geography has recently made progress : 理论和定量地理学最近取得了进展
% It includes simultaneously geographical theories, complex models of simulation and empirical analysis : 它同时包括地理理论，模拟的复杂模型和经验分析
% We build on this legacy 我们建立在这个遗产上
% we study relations between territories and networks 我们研究领土和网络之间的关系
% Our contribution consists in two distinct parts 我们的贡献包括两个不同的部分
% First we develop a new theory of territorial systems 首先，我们开发一个新的领土系统理论
% It emphasizes on the role of networks in coupled evolution processes 它强调网络在耦合演化过程中的作用
% we build on several modeling and empirical previous contributions 我们建立在几个建模和经验的以前的贡献
% , in order to speculate the following hypothesis. 以便推测以下假设
%  within Pumain's Evolutive Urban Theory 在Pumain的演化城市理论
% the existence of subsystems evolving together is equivalent to the existence of morphogenetic processes in which networks are significant drivers  一起演化的子系统的存在等同于形态发生过程的存在，其中网络是重要的驱动因素
% Implications include the necessity of networks in explaining territorial systems dynamics 影响包括网络在解释领土系统动力学中的必要性
% but also a modular decomposition of these systems into local stationary processes in space or time. 而且还将这些系统模块化地分解成空间或时间中的局部固定过程。
% In a second part, we discuss practical implications of the theory on a specific example. 在第二部分中，我们讨论该理论对具体示例的实际含义
% simulates the co-evolution of land-use and transportation infrastructure within a Mega-city Region
%
%

Recent advances in Theoretical and Quantitative Geography have witnessed the co-construction of geographical theories, complex models of simulation and Empirical analysis. We build on this legacy to shed light on relations between territories and networks. Our contribution consists in two distincts parts. First we develop a new theory of territorial systems, that emphasizes on the role of networks in co-evolutive processes. More precisely, we speculate, building on several modeling and empirical previous contributions, that within Pumain's Evolutive Urban Theory~\cite{pumain1997pour}, the existence of co-evolutive subsystems is equivalent to the existence of morphogenetic processes in which networks are significant drivers. Implications include the necessity of networks in explaining territorial systems dynamics, but also a modular decomposition of these systems into local stationary processes in space or time.

In a second part, we discuss practical implications of the theory through the adaptation of a agent-based model introduced in~\cite{le2015modeling}, which simulates the co-evolution of land-use and transportation infrastructure within a Mega-city Region. In particular, it includes game theoretical modeling of decision making processes for transportation governance. The model is partially validated on synthetic data by retrieving expected stylized facts on emergent infrastructure. We propose then to apply it to the Mega-city Region of Pearl River Delta, China, which is very typical of included characteristics such as the presence of a strong competition between inner cities and the recent realization, construction or planning of numerous large-scale transportation infrastructures. Calibration of the model will allow first to infer information on potentially hidden governance processes when applied on real or planned infrastructure, and secondly to compare among different possible governance frameworks when applied on hypothetical optimized infrastructure.



\bigskip
\bigskip

%%%%%%%%%%%%%%%%%%%%
%% Biblio
%%%%%%%%%%%%%%%%%%%%


%\begin{multicols}{2}

%\setstretch{0.3}
%\setlength{\parskip}{-0.4em}


\bibliographystyle{apalike}
\bibliography{biblio}

%\end{multicols}

\end{document}
